---
title: "Demonstration of Central limit Theorem"
author: "Gowtham Talluru"
date: "July 12, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Brief about Central Limit Theorem (CLT)
- Named after Russian Mathematician Aleksandr Lyapunov.
- CLT states that sum of independent and identically distributed random variables with
  finite variances will tend to a normal distribution as the number of variables grows.

## Methodology

- The Population Size, mean and standard deviation are taken Input
- A Normal distribution is generated using above inputs using rnorm function
- The slider can interactively get the value of sample size
- The mean of 100 random sample sizes are plotted in bottom histogram to demonstrate 
  central limit theorem

## Sample Code

```{r cars, echo = TRUE, message=FALSE, fig.show= 'hide'}
    x <- rnorm(1000, 100, 20)  # Population Distribution

    z<-data.frame()   # Sample mean data frame
    for(i in 1:100){
    y <- sample(1:1000, 30, replace = F)
    sampleMean<- mean(x[y])
    z[i,1] <- sampleMean
        hist(x, main="Population distribution of mean=100 and sd=100", 
         col = 'red',
         xlim=c(40, 160))

    hist(z[,1], main="Sample mean Distribution  n=30", col = 'blue',
         xlim=c(40, 160))
    }

```

## Population and Sample Distribution

```{r plots, echo=FALSE}
    par(mfrow=c(1,2))
    hist(x, main="Population of mean=100 and sd=100", 
         col = 'red',
         xlim=c(40, 160))

    hist(z[,1], main="Sample mean Distribution  n=30", col = 'blue',
         xlim=c(40, 160), xlab="Sample Mean")

```


## Application
- Plotting both population and sample distribution will provide a qualitative means to conduct hypothesis testing.
- The above plots will also help in qualitatively determining sample sizes for various studies based on population parameters and required confidence level.

# Thank You