---
title: "ML"
author: "Gowtham Talluru"
date: "May 27, 2017"
output: html_document
---
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)
library(ggplot2)
library(reshape2)
library(rpart)
library(party)                          # for visualizing trees
library(partykit)                       # for visualizing trees
library(rattle)      		                # fancy tree plot
library(rpart.plot)	 
library(caret)
library(randomForest)
```
1. Initial data provide is processed and predictors with null values and NAs were removed. Similar predictors in
test data were also removed.
2. Predictors which doesnot have any significance in the prediction such as "Time Stamp", "ID NO", " Name" etc were
dropped from the data

```{r, message=TRUE, warning=FALSE, include=TRUE}
Processed_data <- read_excel("D:/Coursera/Data Science/Practical machine learning/Pml/Processed data.xlsx", 
     sheet = "Train-final")

Processed_data$Slno<-NULL
Processed_data$user_name<-NULL
Processed_data$Data<-NULL
Processed_data$classe<-as.factor(Processed_data$classe)
```
3. 20% test data set is seperated to evaluate the model built using Random forests
```{r}
train<-createDataPartition(Processed_data$classe, p=0.80,list=FALSE)

Train_Data<-Processed_data[train,]
Test_Data<-Processed_data[-train,]

table(Train_Data$classe)
table(Test_Data$classe)
```

4. Random forests model is selected because of its advantages such as feature selection, resistence to over fitting and ability to handle mulitcollinearity
5. Random forests model is built using 500 trees and "m" (number of random predictors at each split) of 15
6. The resultant confusion matrix using the seperated 20% test data is shown below.

```{r, message=TRUE, warning=FALSE, include=TRUE}

RF_model<-randomForest(classe~., data=Train_Data, ntree=500, mtry=15)
p_RF<-predict(RF_model, newdata=Test_Data)

confusionMatrix(p_RF, Test_Data$classe)
```

7. The classification of the 20 test instances are as follows

```{r, message= TRUE, include = TRUE}
finalTest <- read_excel("D:/Coursera/Data Science/Practical machine learning/Pml/Processed data.xlsx",
sheet = "Test")

testResult<-predict(RF_model, newdata=finalTest)
testResult
```




