---
title: "Regression models"
author: "Gowtham Talluru"
date: "January 21, 2017"
output: word_document
---
##Executive Summary:
The distibution of mpg between manual and auto transmission is shown in figure-1. By priliminary observation it can be observed that manul has more mileage than auto transmission. The statistical significance of this difference will be verified by the regression models.Three different models 
i. model_nonscaled
ii. model_scaled 
iii. model_sign_var
each with original attributes, scaled attributes and selected variables. It can be observed from the models that manual transmission vehicles has 2.9 mpg higher than similar automatic transmission vehicle but the p value of this coefficient is higher than 0.19 which indicates that it is not statistically significant.

##Coefficients
The coefficint for am(1-manual, 0-auto) is 2.9. It means manual transmission car will 
have 2.9mpg higher than exactly similar automatic transmission car.

##Models
Three models are fitted
i. model_nonscaled
ii. model_scaled 
iii. model_sign_var
model_nonscaled uses original attributes whereas model_scaled used scaled and centered variables. model_sign_var uses variables with significance higher than certain limit.

##Question of Interest
It can be observed from the models that manual transmission vehicles has 2.9 mpg higher than similar automatic transmission vehicle but the p value of this coefficient is higher than 0.19 which indicates that it is not statistically significant.

##Residual plot and diagnostics
No significant trends are observed from the residual plots which indicates that the models
are good.

##Uncertainty
The p value of beta realted to transimission is 0.19. It means that the probability of making
a type-1 error with this estimate is 19%. 




```{r cars, message=FALSE, warning=FALSE}
library(glmnet)
library(caret)
data("mtcars")

str(mtcars)

#Comparision using boxplot
boxplot(mpg~am, data=mtcars, main="Distribution of mpg", ylab="mpg", xlab=" 0-auto, 1-manual", col=c("red","blue"))

# Converting numbers into factors
mtcars$cyl<-as.factor(mtcars$cyl)
mtcars$vs<-as.factor(mtcars$vs)
mtcars$am<-as.factor(mtcars$am)
mtcars$gear<-as.factor(mtcars$gear)
mtcars$carb<-as.factor(mtcars$vs)

#Centering and scaling the numeric parameters
mtcars_scaled<-mtcars
mtcars_scaled$disp<-(mtcars_scaled$disp-mean(mtcars_scaled$disp))/sd(mtcars_scaled$disp)
mtcars_scaled$hp<-(mtcars_scaled$hp-mean(mtcars_scaled$hp))/sd(mtcars_scaled$hp)
mtcars_scaled$drat<-(mtcars_scaled$drat-mean(mtcars_scaled$drat))/sd(mtcars_scaled$drat)
mtcars_scaled$wt<-(mtcars_scaled$wt-mean(mtcars_scaled$wt))/sd(mtcars_scaled$wt)
mtcars_scaled$qsec<-(mtcars_scaled$qsec-mean(mtcars_scaled$qsec))/sd(mtcars_scaled$qsec)




```


```{r quant}
#Initial model without centering and scaling the data
model_nonscaled<-lm(mpg~., data=mtcars)
summary(model_nonscaled)
par(mfrow=c(2,2))
plot(model_nonscaled)

model_scaled<-lm(mpg~., data = mtcars_scaled)
model_sign_var<-lm(mpg~hp+wt+am+qsec+vs, data=mtcars_scaled)

```

```{r}
summary(model_scaled)
summary(model_sign_var)

```

