---
title: "Basic Inferential data Analysis"
author: "Gowtham Talluru"
date: "January 22, 2017"
output: word_document
---

# 2. Inferential Data Analysis

## 2.1 Exploratory data Analysis
```{r tooth growth}
data("ToothGrowth")
summary(ToothGrowth)
par(mfrow=c(1,2))
boxplot(len~dose, data=ToothGrowth[ToothGrowth$supp=="OJ",],
        main="Orange juice- Tooth growth", colour=c("red","green","blue"))

boxplot(len~dose, data=ToothGrowth[ToothGrowth$supp=="VC",],
        main="Ascorbic acid- Tooth growth", colour=c("red","green","blue"))
```

## 2.2: Hypothesis test
```{r}
#Comparing 0.5 and 1 dose of supp VC
#Seperating each group
VC_0.5<-ToothGrowth$len[((ToothGrowth$supp=="VC")&(ToothGrowth$dose==0.5))]
VC_1<-ToothGrowth$len[((ToothGrowth$supp=="VC")&(ToothGrowth$dose==1))]

#Comparing variance
var.test(VC_0.5,VC_1)

#Since p value is greater than 0.05 homogeneous variance can be assumed
#t test for both groups

t.test(VC_0.5,VC_1, var.equal = TRUE, paired = FALSE)

```
Since the p value is very small, it can be infered that there is significant 
change in toothgrowth with increase in dose


```{r}
#Comparing 0.5 dose of supp OJ and VC
#Seperating each group
VC_0.5<-ToothGrowth$len[((ToothGrowth$supp=="VC")&(ToothGrowth$dose==0.5))]
OJ_0.5<-ToothGrowth$len[((ToothGrowth$supp=="OJ")&(ToothGrowth$dose==0.5))]

#Since p value is greater than 0.05 homogeneous variance can be assumed
#t test for both groups

t.test(VC_0.5,OJ_0.5, var.equal = TRUE, paired = FALSE)

```


## 2.3: Conclusion
From the above two hypothesis testing it can be infered that there is statistically significant difference in tooth growth between 0.5 and 1 dose of VC(ascorbic acid) and 0.5 dose of OJ(orange juice) and VC( ascorbic acid). The difference in tooth growth of other groups can be infered similarly.




```{r}

library(rattle)
data(segmentationOriginal)


set.seed(125)
train<-createDataPartition(segmentationOriginal$Case, p=0.8, list=FALSE)


trainset<-segmentationOriginal[train,]
testset<-segmentationOriginal[-train,]

model<-train(Case~., data=trainset,method="rpart")


model<-rpart(Case~., data=trainset)


fancyRpartPlot(model)
```




